Human capital mediates natural selection in contemporary humans

### 2021-07-22 why-natural-selection.Rmd
Report results from within-siblings regressions

```{r calc-res-earn-educ}

res_ee_control %<>% filter(term == score_name)
n_05 <- sum(res_ee_control$p.value < 0.05)
n_05_33 <- sum(res_ee_control$p.value < 0.05/33)

@@

* We estimate the causal effect of each PGS on fertility using within-sibling
  regressions.
  - These are causal due to lottery of meiosis
  - Also do the fertility PGS

* We also estimate the effect of PSEA and fertility PGS on education and
  earnings, again within-siblings
  - confirm PSEA affects both; does fertility? What if we put both in together?
* Lastly we again estimate effect of PSEAs on fertility, 
  controlling for education and earnings in first job. 
  - We cannot exclude some reverse causality from EA3 -> fertility -> education/income
  - Our theory indeed assumes that people simultaneously choose N and s, while taking into
    account their effect on Y. We don't make claims about exogenous (chance) variation
    in Y or in N. But note that exog variation in N1 (i.e. no choice, treat
    it as fixed to different values) would affect s. OTOH, if EA3 affects $a$ and
    thereby changes things, you don't get predictions 2 or 4.

* Results:
  - in bivariate within-sib regs, extraversion, EA2/3, ADHD and some others are
    weakly significant (only at .05!)
  - adding education as a control barely changes effect sizes
  - adding income as a control increases effect sizes (of EA at least)
  - by contrast, in pooled regs, adding education shrinks effect sizes, and
    adding income does nothing.
  - relevant: effect of EA2 on education is halved within-sib, and effect of
    education on fertility is halved within-sib
  - whereas, effect of EA2 on income is halved within-sib, but effect of income
    on fertility is multiplied by 10!! within-sib. Much of this is because
    of year of birth confound i.e. declining fertility and also retired people
    have smaller income

@@

>>>>

@@

```{r calc-fe-fertility}
drake::loadd(res_fe_fertility)
drake::loadd(res_all)

res_all %<>% filter(dep.var == "N children", reg.type == "controlled")
res_fe_fertility %<>% filter(Regression == "Raw")
res_all %<>% left_join(res_fe_fertility, by = "term")

cor_pool_within <- cor.test(res_all$estimate.x, res_all$estimate.y)
```

One interpretation of these results is that natural selection on polygenic
scores is driven by their correlation with human capital, which affects
fertility as our model describes. This does not necessarily imply that polygenic
scores are causal of either human capital or fertility. Instead they may simply
correlate with different family environments, for example via "genetic nurture"
effects [@kong2018nature] or population stratification. To test whether
polygenic scores cause fertility, we estimate equation \@ref(eq:regression)
among full siblings, controlling for sibling group fixed effects (Appendix
Figure \@ref(fig:plot-res-fe-fertility)). Results are never significant after
Bonferroni correction. The correlation between pooled and within-sibling effect
sizes is `r cor_pool_within$estimate` (p = `r cor_pool_within$p.value`), and within-sibling effects are about 1/3 smaller. Thus, there is only weak evidence for causal
links from polygenic scores to fertility, with at least some of the effect
coming from gene-environment correlation. If we re-estimate the within-sibling
regressions adding a control for education, effect sizes barely change. This
suggests that our measures of education only mediate between-sibling-group
differences in fertility; among siblings, other causal channels are more
important.

* Results:
  - in bivariate within-sib regs, extraversion, EA2/3, ADHD and some others are
    weakly significant (only at .05!)
  - adding education as a control barely changes effect sizes
  - adding income as a control increases effect sizes (of EA at least)
  - by contrast, in pooled regs, adding education shrinks effect sizes, and
    adding income does nothing.
  - relevant: effect of EA2 on education is halved within-sib, and effect of
    education on fertility is halved within-sib
  - whereas, effect of EA2 on income is halved within-sib, but effect of income
    on fertility is multiplied by 10!! within-sib. Much of this is because
    of year of birth confound i.e. declining fertility and also retired people
    have smaller income
  - within-sib and pooled coefs for PGS are highly correlated; within coefs
    are about 67% of the size.

* Interpretation:
  - some families are better educated and have fewer children
  - they also have selected-against PG scores like what we see
  - scores aren't necessarily causal for children, they may be
    associated
  - the between-family effect goes through education, but not the 
    within-family effect (cf. the paper on causality)

@@

\FloatBarrier
\clearpage
## Within-siblings regressions

```{r plot-res-fe-fertility, fig.cap = "Selection effects controlling for sibling-group fixed effects. \"Controlled\" regressions add a control for education (left education before 16, 16-18, or after 18)."}

drake::loadd(res_fe_fertility)

standard_ggplot(res_fe_fertility, fill_col = Regression, n_regs = 33, fill_direction = -1)
